```{r echo=FALSE, cache.lazy=FALSE}
echocode=F
scriptname="graphs.Rmd"
setwd("~/Skinner/Papers/Working/NoSplitting/Significance/newSoft")
load('simdata.rda')
load('experiment1.rda')
myexper<-info[[1]]
library(simCity)
rho<-unlist(coefs.unique(myexper)[1])
rel<-unlist(coefs.unique(myexper)[2])

```

Simulation for Significance of splitting (Jan. 2014)
========================================================

Based on Jamie's notes
[go to pdf ](https://www.dropbox.com/s/jtw5z4uyrxj5rcg/Notes%202012-12-21.pdf)


## Setup
### Model
Two continuous latent variables (\(\eta\) and \(\xi\) ) are created with <b>N</b> cases, sharing a correlation equal to \(\rho\). A measure \(x\) of \(\xi\) is created with reliability \(rel\), and then  is dichotomized accordingly to \(p\) \(1-p\) into \(c\). The correlations \( r_pe=r(\eta,x) \)  and \( r_pb=r(\eta,c) \) are computed, their p-value and significance (at .05) is recorded.
### Design

_N_={`r myexper$N `}
<br>
_rho_={`r rho[1] `}
<br>
_rel_={`r rel `}

each combination of _N_ and _rho_ is repeated 100 times.


#### Proportions of samples where both correlations are significant (_both_), only the .biserial (_Split_), and only the Pearson (_continuos_) by reliability 


```{r fig.width=8, fig.height=6, echo=echocode,warning=echocode, results='hide', comment=''}
md<-function(rho,p1,p2) {
s0<-simdata[simdata$rho==rho,]
sums<-sum(s0$x)
ss1<-aggregate(s0$x,list(s0$rel),sum)
s1<-aggregate(s0,list(s0$rel,s0$sigpe,s0$sigpb),sum)
s2<-s1[(s1$Group.2==p1 & s1$Group.3==p2),]
s2$prop<-s2$x/ss1[,2]
s2
}
dat<-md(0,1,0)
plot(dat$prop~dat$Group.1,ylab='prop. of runs',xlab='rel',pch=21,type='b',ylim=c(0,.05))
dat<-md(0,0,1)
points(dat$prop~dat$Group.1,pch=25,bg=0,type='b')
dat<-md(0,1,1)
points(dat$prop~dat$Group.1,pch=19,bg=0,type='b')
legend('topright',inset=c(-.05,-.1),
       c('Continuous.','Split','Both'),
       pch=c(21,25,19),
       xpd=T,
       pt.bg=c(0,1),
       title='Indicator',
      box.col = "black",bg = "white")

```


#### Proportion of significant results by N broken down by indicator type


```{r fig.width=8, fig.height=6, echo=echocode,warning=echocode, results='hide', comment=''}

md<-function(rho,p1,p2) {
s0<-simdata[simdata$rho==rho,]
sums<-sum(s0$x)
ss1<-aggregate(s0$x,list(s0$ncase),sum)
s1<-aggregate(s0,list(s0$ncase,s0$sigpe,s0$sigpb),sum)
s2<-s1[(s1$Group.2==p1 & s1$Group.3==p2),]
s2$prop<-s2$x/ss1[,2]
s2
}
dat<-md(0,1,0)
plot(dat$prop~dat$Group.1,ylab='prop. of runs',xlab='N',pch=21,type='b',ylim=c(0,.05))
dat<-md(0,0,1)
points(dat$prop~dat$Group.1,pch=25,bg=0,type='b')
dat<-md(0,1,1)
points(dat$prop~dat$Group.1,pch=19,bg=0,type='b')


legend('topright',inset=c(-.05,-.1),
       c('Continuous.','Split','Both'),
       pch=c(21,25,19),
       xpd=T,
       pt.bg=c(0,1),
       title='Indicator',
      box.col = "black",bg = "white")


```


#### Mean abs distance between Pearson correlation and \(\rho\) results by reliability broken down by indicator significance

<i>in this case mean distance=effect size in terms of r</i>

```{r fig.width=8, fig.height=6, echo=echocode,warning=echocode, results='hide', comment=''}
md<-function(rho,p1,p2) {
s0<-simdata[simdata$rho==rho,]
sums<-sum(s0$x)
ss1<-aggregate(s0$x,list(s0$rel),mean)
s1<-aggregate(s0,list(s0$rel,s0$sigpe,s0$sigpb),mean)
s2<-s1[(s1$Group.2==p1 & s1$Group.3==p2),]
s2$prop<-s2$x/ss1[,2]
s2
}
dat<-md(0,1,0)
plot(abs(dat$rpe.cor)~dat$Group.1,ylab='delta',xlab='rel',pch=21,type='b',ylim=c(-.2,.2))
dat<-md(0,0,1)
points(abs(dat$rpe.cor)~dat$Group.1,pch=25,bg=0,type='b')
dat<-md(0,1,1)
points(abs(dat$rpe.cor)~dat$Group.1,pch=19,bg=0,type='b')


legend('topright',inset=c(-.05,-.1),
       c('Continuous.','Split','Both'),
       pch=c(21,25,19),
       xpd=T,
       pt.bg=c(0,0,1),
       title='Which significant?',
      box.col = "black",bg = "white")

```

#### Mean abs distance between .biserial correlation and \(\rho\) results by reliability broken down by indicator significance

<i>in this case mean distance=effect size in terms of r</i>

```{r fig.width=8, fig.height=6, echo=echocode,warning=echocode, results='hide', comment=''}
md<-function(rho,p1,p2) {
s0<-simdata[simdata$rho==rho,]
sums<-sum(s0$x)
ss1<-aggregate(s0$x,list(s0$rel),mean)
s1<-aggregate(s0,list(s0$rel,s0$sigpe,s0$sigpb),mean)
s2<-s1[(s1$Group.2==p1 & s1$Group.3==p2),]
s2$prop<-s2$x/ss1[,2]
s2
}
dat<-md(0,1,0)
plot(abs(dat$rpe.cor)~dat$Group.1,ylab='delta',xlab='rel',pch=21,type='b',ylim=c(-.2,.2))

dat<-md(0,0,1)
points(abs(dat$rpb.cor)~dat$Group.1,pch=25,bg=0,type='b')
dat<-md(0,1,1)
points(abs(dat$rpb.cor)~dat$Group.1,pch=19,bg=0,type='b')

legend('topright',inset=c(-.05,-.1),
       c('Continuous.','Split','Both'),
       pch=c(21,25,19),
       xpd=T,
       pt.bg=c(0,0,1),
       title='Which significant?',
      box.col = "black",bg = "white")

```

############## not null hypothesis #############

### Design

_N_={`r myexper$N `}
<br>
_rho_={`r rho[2] `}
<br>
_rel_={`r rel `}

each combination of _N_ and _rho_ is repeated 100 times.


#### Proportions of samples where both correlations are significant (_both_), only the .biserial (_Split_), and only the Pearson (_continuos_) by reliability 


```{r fig.width=8, fig.height=6, echo=echocode,warning=echocode, results='hide', comment=''}
md<-function(rho,p1,p2) {
s0<-simdata[simdata$rho==rho,]
sums<-sum(s0$x)
ss1<-aggregate(s0$x,list(s0$rel),sum)
s1<-aggregate(s0,list(s0$rel,s0$sigpe,s0$sigpb),sum)
s2<-s1[(s1$Group.2==p1 & s1$Group.3==p2),]
s2$prop<-s2$x/ss1[,2]
s2
}
dat<-md(.5,1,0)
plot(dat$prop~dat$Group.1,ylab='prop. of runs',xlab='rel',pch=21,type='b',ylim=c(0,1))
dat<-md(.5,0,1)
points(dat$prop~dat$Group.1,pch=25,bg=0,type='b')
dat<-md(.5,1,1)
points(dat$prop~dat$Group.1,pch=19,bg=0,type='b')
legend('topright',inset=c(-.05,-.1),
       c('Continuous.','Split','Both'),
       pch=c(21,25,19),
       xpd=T,
       pt.bg=c(0,0,1),
       title='Indicator',
      box.col = "black",bg = "white")

```


#### Proportion of significant results by N broken down by which indicator is significant


```{r fig.width=8, fig.height=6, echo=echocode,warning=echocode, results='hide', comment=''}

md<-function(rho,p1,p2) {
s0<-simdata[simdata$rho==rho,]
sums<-sum(s0$x)
ss1<-aggregate(s0$x,list(s0$ncase),sum)
s1<-aggregate(s0,list(s0$ncase,s0$sigpe,s0$sigpb),sum)
s2<-s1[(s1$Group.2==p1 & s1$Group.3==p2),]
s2$prop<-s2$x/ss1[,2]
s2
}
dat<-md(.5,1,0)
plot(dat$prop~dat$Group.1,ylab='prop. of runs',xlab='rel',pch=21,type='b',ylim=c(0,1))
dat<-md(.5,0,1)
points(dat$prop~dat$Group.1,pch=25,bg=0,type='b')
dat<-md(.5,1,1)
points(dat$prop~dat$Group.1,pch=19,bg=0,type='b')

legend('topright',inset=c(-.05,-.1),
       c('Continuous.','Split','Both'),
       pch=c(21,25,19),
       xpd=T,
       pt.bg=c(0,1),
       title='Indicator',
      box.col = "black",bg = "white")


```


#### Mean abs distance between Pearson correlation and \(\rho\) results by reliability broken down by indicator significance


```{r fig.width=8, fig.height=6, echo=echocode,warning=echocode, results='hide', comment=''}
rho<-.5

md<-function(rho,p1,p2) {
s0<-simdata[simdata$rho==rho,]
sums<-sum(s0$x)
ss1<-aggregate(s0$x,list(s0$rel),mean)
s1<-aggregate(s0,list(s0$rel,s0$sigpe,s0$sigpb),mean)
s2<-s1[(s1$Group.2==p1 & s1$Group.3==p2),]
s2$prop<-s2$x/ss1[,2]
s2
}
dat<-md(.5,1,0)
plot(abs(dat$rpe.cor-rho)~dat$Group.1,ylab='delta',xlab='N',pch=21,type='b',ylim=c(0,.5))
dat<-md(.5,0,1)
points(abs(dat$rpe.cor-rho)~dat$Group.1,pch=25,bg=0,type='b')
dat<-md(.5,1,1)
points(abs(dat$rpe.cor-rho)~dat$Group.1,pch=19,bg=0,type='b')


legend('topright',inset=c(-.05,-.1),
       c('Continuous.','Split','Both'),
       pch=c(21,25,19),
       xpd=T,
       pt.bg=c(0,0,1),
       title='Which significant?',
      box.col = "black",bg = "white")

```

#### Mean abs distance between .biserial correlation and \(\rho\) results by reliability broken down by indicator significance


```{r fig.width=8, fig.height=6, echo=echocode,warning=echocode, results='hide', comment=''}
rho<-.5
md<-function(rho,p1,p2) {
s0<-simdata[simdata$rho==rho,]
sums<-sum(s0$x)
ss1<-aggregate(s0$x,list(s0$rel),mean)
s1<-aggregate(s0,list(s0$rel,s0$sigpe,s0$sigpb),mean)
s2<-s1[(s1$Group.2==p1 & s1$Group.3==p2),]
s2$prop<-s2$x/ss1[,2]
s2
}
dat<-md(rho,1,0)
plot(abs(dat$rpe.cor-rho)~dat$Group.1,ylab='delta',xlab='rel',pch=21,type='b',ylim=c(0,.5))

dat<-md(.5,0,1)
points(abs(dat$rpb.cor-rho)~dat$Group.1,pch=25,bg=0,type='b')
dat<-md(.5,1,1)
points(abs(dat$rpb.cor-rho)~dat$Group.1,pch=19,bg=0,type='b')

legend('topright',inset=c(-.05,-.1),
       c('Continuous.','Split','Both'),
       pch=c(21,25,19),
       xpd=T,
       pt.bg=c(0,0,1),
       title='Which significant?',
      box.col = "black",bg = "white")



```

Proportions of samples when each of the correlation is significant



```{r fig.width=8, fig.height=6, echo=echocode,warning=echocode, comment=''}

result<-simdata$sigpe+10*simdata$sigpb
s1<-aggregate(simdata$x,list(simdata$sigpe,simdata$sigpb,simdata$rho),sum)
s1
ss<-sum(simdata$x)
s1$prop<-s1$x/(ss*2)
names(s1)<-c("conti.sig","split.sig","rho","freq","prop")
print(s1)
cc<-s1[s1[,1]==1,]
a1<-cc[c(1,2),]
a2<-cc[c(3,4),]
a1
a2
```

